---
title: "hw02-jade-wang"
output: github_document
---
```{r}
library(readr)

```
* Importing the Data in R
     + First method: read_csv()
     + Second method: read.csv()
     
```{r}
table1 <- read_csv('/Users/Jade/nba2017-player-statistics.csv'
                   , col_types =cols(Position = col_factor(levels = NULL)))

table1
```
```{r}
table2 <- read.csv('/Users/Jade/nba2017-player-statistics.csv', 
                   stringsAsFactors = FALSE,
                   colClasses = c( Position = 'factor'))
table2
```

* Processing the Data

```{R}
dat <- "/Users/Jade/nba2017-player-statistics.csv"
df <- read.csv("/Users/Jade/nba2017-player-statistics.csv")

df$Experience <- as.character(df$Experience)
experience1 <- replace(df$Experience , df$Experience == 'R' , 0)

df$Experience <- as.integer(experience1)
table3 <- df
table3

```

* Performance of Players
```{R}
missed_FG <- df$FGA - df$FGM
missed_FT <- df$FTA - df$FTM
PTS <- 2*df$Points2 + 3*df$Points3 + 1*df$FTM
REB <- df$OREB + df$DREB
MPG <- df$MIN / df$GP
EFF <- (PTS + REB + df$AST + df$STL + df$BLK - missed_FG - missed_FT - df$TO)/df$GP


table4 <- data.frame(
  'missed field goals' = missed_FG,
  'missed free throws' = missed_FT,
  'total points' = PTS,
  'total rebounds' = REB,
  'minutes per game' = MPG,
  'efficiency' = EFF)
table4

summary(EFF)
hist(EFF)
```
```{r}
# Display names, teams, salary and EFF of the top-10 players
library(dplyr)
df1 <- mutate(df, EFF)
df2 <- arrange(df1, EFF)
top10_EFF <- slice(df2, 1:10)
select(top10_EFF, Player, Team, Salary, EFF)
```

```{r}
# names of player with negative EFF
negative_EFF <- filter(df1, EFF < 0)
select(negative_EFF, Player)
```


```{r}
df <- read.csv("/Users/Jade/nba2017-player-statistics.csv")

missed_FG <- df$FGA - df$FGM
missed_FT <- df$FTA - df$FTM
PTS <- 2*df$Points2 + 3*df$Points3 + 1*df$FTM
REB <- df$OREB + df$DREB
MPG <- df$MIN / df$GP
EFF <- (PTS + REB + df$AST + df$STL + df$BLK - missed_FG - missed_FT - df$TO)/df$GP
df5 <- data.frame(PTS, REB, df$AST, df$STL, df$BLK,- missed_FG, - missed_FT,  - df$TO)
list5 <- c(PTS, REB, df$AST, df$STL, df$BLK,- missed_FG, - missed_FT,  - df$TO)


#correlation between EFF and other variables
corr <- c(cor(EFF,PTS), cor(EFF,REB), cor(EFF,df$STL), cor(EFF,df$BLK), cor(EFF,- missed_FG), cor(EFF,- missed_FT), cor(EFF,- df$TO))

corr <- sort(corr, decreasing = TRUE)
corr

#barchart in decreasing order
list6 <- c('PTS','REB','STL','AST','BLK','Missed_FT', 'Missed_PG', 'TO')
barplot(corr, main = 'Correlations between Player Stats and EFF')


plot1 <- plot(x = EFF, y = df$Salary, main = 'EFF and Salary', xlab = 'Efficiency', ylab = 'Salary' ) #scatterplot of EFF and salary
lines(lowess(EFF, df$Salary)) #adding lowess line

#From the scatterplot, we can find a generally positive correlation between players' efficiency and salary.

DF <- cbind(df,EFF,PTS,REB,MPG, missed_FG, missed_FT)

players2 <- filter(DF, MPG >= 20)

plot(players2$EFF, players2$Salary, main = 'EFF and Salary of more established players', xlab = 'Efficiency', ylab = 'Salary')
lines(lowess(players2$EFF, players2$Salary))

df6 <- data.frame(players2$PTS, players2$REB, players2$AST, players2$STL, players2$BLK, players2$missed_FG, players2$missed_FT, players2$TO)

cor(players2$EFF, df6) #correlations

# I notice that more established players tend to have a less strong positive correlation between EFF and Salary. I think, compared with new players, those established players' salaries would also depend on their popularity, reputation and social influence, rather than merely their efficiency during the season.

```

* Comments and Reflections
     + I spent a lot of time on processing the data, like creating the data frame.
     + Plotting is not too hard as long as I know which function to use with.
     + I spent 5+ hrs on this homwwork, and almost half time was on searching for proper functions to use.
     + I still feel confused by the class of data, like difference between character, factor and vector.
     + What makes me feel most proud of is I finally finish this homework and I do learn a lot.
       
