---
title: "Post02: A quick guide to Shiny apps"
subtitle: "Jade Wang 11/30/2017"
output: html_document
Arthor: Jade Wang
Date: 11/30/2017
runtime: shiny
---


# Introduction & Background

### Shiny enables us to write powerful interactive web applications entirely in R. Using R, we can create a user interface and server and Shiny compiles our code into the HTML, CSS and JavaScript needed to display our app online.


### From my own perspective, Shiny is one of the most challenging topics in this course, so I want to make some breif guidelines about how to do Shiny app in a simple way, and I hope it will help! Shiny may be hard to handle at first, but I hope my post could give everyone a more clear understanding of how Shiny works. 

## Let's nail it!

<img src = "/Users/Jade/stat133-hws-fall17/post02/images/bb1.jpg" />


## *Preparation*

#### Let's start by loading some pakages first.

#### In order to run Shiny and follow the code on this post you should make sure you have RStudio software and the shiny R package.
```{r}
library(shiny)
```


# The Most Basic Shiny App

## In its simplest form, a Shiny application requires a server function to do the calculations and a user interface. 

### In the example, we will put only some text in UI, so the app may look strange, but it is a real Shiny app.

#### The code is as below:
```{r}
library(shiny)

server <- function(input, output, session) { } # the server

ui <- basicPage(
'Here is a Shiny app at simplest.'
 ) # the user interface

shinyApp(ui = ui, server = server) # it runs the app
```

### Yes, I know the simplest Shiny app looks ugly and strange, but I promise things will get much better as we add more elements into our Shiny app.

<img src = "/Users/Jade/stat133-hws-fall17/post02/images/bb3.png" />

# Set up your user interface (UI)

## *1. Layout your unser interface*

#### When we run the function _shinyApp()_ our R code is essentially compiled to web-friendly languages HTML, JavaScript and CSS. We can take advantage of Shiny page template functions to layout our app.

#### In this example, we will focus on the simplest function, by creating sidebarPanel and mainPanel.

#### The code is as below:
```{r}
#Example app: basic user interface
library(shiny)

ui <- fluidPage(
  titlePanel("This is a Shiny app with the simplest layout"),
  sidebarLayout(
    sidebarPanel(
      'Sidebar'
    ),
    mainPanel(
      'Main panel'
    )
  )
)

server <- function(input, output, session) { } # the server

shinyApp(ui, server)

```

## *2. Adding elements with tags$xxx*

#### It's easy to add HTML elements to our Shiny app using HTML tags.
#### There are more than 100 HTML elements available.

#### In the example, we add HTML tags using _tags$blockquote_, and common elements like _h1_ and _h3_
#### The code is as below:
```{r}
#Example app: using HTML tags in Shiny
library(shiny)

ui <- fluidPage(
  h1('Here is H1 without tags$'),
  tags$blockquote('The block quote here requires tags$'),
  h3('Here is H3 without tags$'),
  code('Here is the data: data.frame(a=1:20, b=1:20)')
)

server <- function(input, output, session) { } # the server

shinyApp(ui, server)
```

## *3. Layout Shiny app using Bootstrap's grid system manually*

#### We can use Bootsrep's system directly by specifying rows and co;umns.

#### In the example, we'll use Boostrap elements like rows and columns to manually layout the app. We'll have 2 rows: one is offset by one column and the other has a button (1 column width) and text (5 column width).

#### The code is as below:
```{r}
# Example app: Manual layout using Bootstrap's grid system

ui <- fluidPage(
  
  fluidRow(
    column(6, offset = 1,
           h1("Here is the title")
       )
  ),
  fluidRow(
    column(1,
           actionButton('Button','click')
   ),
   column(5,
          p("Row 2, Column 2 (buttion is col 1)")
          )
  )
  
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```


## *4. Style our user interface (UI) with exsiting themes*

#### We've discussed some approaches to style our Shiny app as above.
#### Still, Bootstrap offers a number of pre-created themes that allow us to style our app with limited coding. 

#### But first, we need to load the shinythemes package from RStudio.
#### The code is as below:
```{r}
# Example app: using a pre-created Bootstrap theme
library(shinythemes)

server <- function(input, output, session) {
  
}

ui <- fluidPage(theme=shinytheme("cosmo"),
  
  titlePanel("App use an existing theme"),
  
  sidebarLayout(
    
    sidebarPanel(
      h3("Note the button has a color of black!!!"),
      actionButton("button", "This is a button")
    ), 
    
    mainPanel(
      tabsetPanel(
        tabPanel("Plot1"), 
        tabPanel("Plot2"), 
        tabPanel("Plot3")
      )
    )
  )
)

shinyApp(ui = ui, server = server)
```


## *6. Add widgets to user interface (sliders, text boxes, etc)*

#### At this point we have scaffolded and added styles to our UI, but we can also add widgets to make UI more interactive, like pull-down menus.

### Note that all widgets must have a unique input ID.

#### In the example, we will add a slider and a text input to our UI.
#### The code is as below:
```{r}
# Example app: allowing inputs

server <- function(input, output, session) {
  
}

ui <- basicPage(
  
  h3("Here are some simple widgets "),
  # a slider
  sliderInput(inputId = "myslider", label = "Here is the label ", min = 0, 
              max = 100, value = c(50, 60)),
  # a text input box
  textInput(
    inputId = "mytext", label = "Text Input", 
    value = "Enter some text here..."
    )
)


shinyApp(ui = ui, server = server)
```


# Now, let's link UI with the server to display our outputs

### In order to return values to the user we need a strategy to:
### (1) grab needed values from the UI in the server
### (2) process as necessary 
### (3) return the result to the UI


## *1. To print text in UI*

#### In the example, we'll use functions of *renderText* and *textOutput* to demonstrate how to print text.

#### The code as below:
```{r}
# Example app: linking UI and the server

server <- function(input, output, session) {
  
  # input$mytext comes from the UI. my_output_text 
  # sent back
  output$my_output_text <- renderText({
    return(input$mytext)
    })
}

ui <- basicPage(
    h3("The value in the text box gets printed below with the *textOutput* function."),
    textInput("mytext", "Input goes here"),
    # my_output_text comes from the server
    
    "Output value is:",
    textOutput("my_output_text")
)

shinyApp(ui = ui, server = server)
```


## *2. Adding a plot to UI*

#### Similiar as the previous example, we'll use functions of *renderPlot* and *plotOutput* to add a plot to UI.

The code as below:
```{r}
server <- function(input, output, session) {
  
  output$my_output_text <- renderText({
    init <- "Input title is: "
    return(paste0(init, input$mytext))
  })
  
  # send plot to the ui as my_output_plot
  output$my_output_plot <- renderPlot({
    plot(1:7, 1:7, pch=16, col=1:7, cex=1:7, main = input$mytext)
  })
  
  
}

ui <- basicPage(
  
  h3("Now we have both text and plot output"),
  textInput("mytext", "Main title goes here"),
  
  # my_output_text comes from the server
  textOutput("my_output_text"),
  # my_output_plot comes from the server
  plotOutput("my_output_plot")
  
)

shinyApp(ui = ui, server = server)


```


## *3. Dynamic UI with functions of  *renderUI* and *OutputUI* *

#### By using *renderUI* and *OutputUI*, we can return a list of UI elements rather than a single object.

#### In the example, we can add new object to the selection list by typing the name in the textbox and then clicking the button.

#### The code as below:
```{r}
# Example app: dynamic UI

server <- function(input, output, session) {
  
  # return a list of UI elements
  output$my_output_UI <- renderUI({
    
    list(
      h4(style = "color:blue;", "My selection list"),
      selectInput(inputId = "myselect", label="", choices = selections)
    )
  })
  
  # initial selections
  selections <- c("New York", "Philadelphia")
  
  # use observe event to notice when the user clicks the button
  # update the selection list. Note the double assignment <<-
  observeEvent(input$mybutton,{
    selections <<- c(input$mytext, selections)
    updateSelectInput(session, "myselect", choices = selections, selected = selections[1])
  })

}

ui <- basicPage(
  
    h3("Using renderUI and uiOutput"),
    uiOutput("my_output_UI"),
    textInput("mytext", ""),
    actionButton("mybutton", "Click to add to selections")

)

shinyApp(ui = ui, server = server)

```

# Finally, let's put some real data into our Shiny app
```{r}
library(shiny)

dat <- read.csv('/Users/Jade/stat133-hws-fall17/post02/data/scores.csv')

ui <- fluidPage(
  titlePanel("Example App"),
  
  h4("Observations"),
  tableOutput("view"),
  
  
  selectInput("select", label = h3("Select box"), 
              choices = list("HW1" = "HW1", "HW2" = "HW2",
                             "HW3" = "HW3", "HW4" = "HW4",
                             "HW5" = "HW5", "HW6" = "HW6",
                             "HW7" = "HW7", "HW8" = "HW8",
                             "HW9" = "HW9"), 
              selected = "HW1"),
  sliderInput("width",
              "Bind width",
              min = 0,
              max = 120,
              value = 50),
  
  selectInput("variableX", label = h3("X-axis variable"),
              choices = list("Test1" = "Test1", "Test2" = "Test2"), 
              selected = "Test1"),
  
  selectInput("variableY", label = h3("Y-axis variable"),
              choices = list("Overall" = "Overall"), 
              selected = "Overall"),
  
  h3("Summary"),
  verbatimTextOutput("summary"),
    
    
    # Show a plot of the generated distribution
    mainPanel(
      tabsetPanel(
        tabPanel("Barplot",  plotOutput("bar")
      )
    )
  )
 )



server <- function(input, output, session) {
  
  output$bar <- renderPlot({
    # generate bins based on input$bins from ui.R
    
    # draw the histogram with the specified number of bins
    grades = factor(dat$Grade, levels = c('A+', 'A', 'A-', 'B+', 'B', 'B-', 'C+', 'C', 'C-', 'D', 'F'))
    barplot(table(grades), col = 1:11, border = 'black')
  })
  
}

shinyApp(ui, server)



```




# Conclusions

### In order to make a dynamic Shiny app, we need to first create user interface (UI) and server, and then link these two parts together to allow inputs and outputs. 

#### From the above examples, we learn that, despite of the most basic structure, we can also style our UI layout with either HTML tags or pre-created Bootstrap themes. 

### Also, in order to link UI with server, we can use functions of *renderUI* and 'OutputUI' to create a dynamic Shiny app.




### I hope everyone has a better idea about Shiny app now! Thank you:)

<img src = "/Users/Jade/stat133-hws-fall17/post02/images/bb2.jpg" />


# References

### Shiny tutorial resources from GitHub
https://shiny.rstudio.com/tutorial/

https://github.com/ucb-stat133/stat133-fall-2017/blob/master/cheat-sheets/shiny-cheatsheet.pdf

https://github.com/ucb-stat133/stat133-fall-2017/blob/master/slides/16-shiny-tutorial.pdf

https://docs.google.com/presentation/d/1_XhkkhXA26izACJyoc5JYrWiF_f96a6tD9ZSI5Tm8Os/pub?start=false&loop=false&delayms=3000&slide=id.gd30d76405_0_29

### Shiny UI layout guide
http://shiny.rstudio.com/articles/layout-guide.html

### HTML tags
http://shiny.rstudio.com/articles/tag-glossary.html

### Bootstrap Basics
https://getbootstrap.com/docs/3.3/examples/grid/








