---
title: "hw03-jade-wang"
output: github_document
---

```{r setup}
library(dplyr)
library(readr)
library(ggplot2)

dat <- read.csv('/Users/Jade/stat133-hws-fall17/hw03/data/nba2017-teams.csv')
```

```{r}
# basic ranking by salary in decreasing order

dat_by_salary <- dat %>%
  select(team, salary) %>%
  arrange(desc(salary)) 
  

ggplot(data = dat_by_salary,
       aes(x = reorder(team, salary, sum), y = salary)) +
  labs(title = 'NBA Teams ranked by Total Salary',
         x = 'Team', y = 'Salary (in millions)') +
  geom_bar(stat = 'identity') +
  coord_flip() +
  geom_hline(aes(yintercept = mean(salary)), col = 'red')



# ranking by total points
dat_by_TP <- dat %>%
  select(team, points) %>%
  arrange(desc(points)) 
  

ggplot(data = dat_by_TP, 
       aes(x = reorder(team, points, sum), y = points)) +
  labs(title = 'NBA Teams ranked by Total Points',
         x = 'Team', y = 'Total Points') +
  geom_bar(stat = 'identity') +
  coord_flip() +
  geom_hline(aes(yintercept = mean(points), col = 'red'))



# ranking by efficiency
dat_by_EFF <- dat %>%
  select(team, efficiency) %>%
  arrange(desc(efficiency)) 
  

ggplot(data = dat_by_EFF, 
       aes(x = reorder(team, efficiency, sum), y = efficiency)) +
  labs(title = 'NBA Teams ranked by Efficiency',
         x = 'Team', y = 'Efficiency') +
  geom_bar(stat = 'identity') +
  coord_flip() +
  geom_hline(aes(yintercept = mean(dat$efficiency), col = 'red'))
```
### From the above three different ranking methods, I find that the top 2 teams in any ranking system would be within three teams: CLE, GSW and LAL. However, the other teams have quite different ranks in each system. Therefore, I think that the strongest teams uauslly remain as the top ones regardless of which ranking system used.

```{R}
# Principal Components Analysis (PCA)

dat <- read.csv('/Users/Jade/stat133-hws-fall17/hw03/data/nba2017-teams.csv')

list1 <- dat %>% select(
  points3_made,
  points2_made,
  points1_made,
  off_rebounds,
  def_rebounds,
  assists,
  steals,
  blocks,
  turnovers,
  fouls)

pca <- prcomp(list1, scale. = TRUE)

#eigenvalues
eigs <- data.frame(
  eigenvalue = pca$sdev^2,
  prop = round(pca$sdev^2/sum(pca$sdev^2) , 4),
  cumprop = round(cumsum(pca$sdev^2/sum(pca$sdev^2)), 4)
)

eigs

df1 <- data.frame(
  pc1 = pca$x[ ,1],
  pc2 = pca$x[ ,2]
)

ggplot(df1, aes(x = pc1, y = pc2)) +
  geom_text(aes(label = dat$team)) +
  geom_hline( yintercept = 0) +
  geom_vline( xintercept = 0)

# Index based on PC1
library(scales)

pc1 <- pca$x[,1]

z1 <- rescale(pc1, to = c(0,100))

s1 <- 100 * (z1 - min(z1))/(max(z1) - min(z1))

df2 <- data.frame(dat$team , s1)

ggplot(data = df2, 
       aes(x = reorder(dat$team, s1, sum), y = s1)) +
  labs(title = 'NBA Teams ranked by scaled PC1',
         x = 'Team', y = 'First PC (scaled from 0 to 100)') +
  geom_bar(stat = 'identity') +
  coord_flip()

```

# Comments and Reflections
    +   I find dealing with different types of data is difficult for me. 
    I have to pay attention to the type of data I'm processing with, or I would not get the results I want.
    +   I find, as long as I collect the correct data frame, 
    plotting becomes much more easier than I first used rmd.
    +   I spent around 6 + hours on this homework,
    but most of the time I was stuck rather than writing codes.
    +   The most time consuming part is re-organizing data base 
    and creating new data frames. Plotting was pretty interesting though.
  

